Bayesian multivariate skew-normal finite mixture model for analysis of infant development trajectories

### 2019-06-04 manuscript/manuscript.tex
approaches

@@

We model the effect of covariates on longitudinal development outcomes through the use of a MSN regression model. The MSN distribution can be represented as the superposition of a MN random variable with a latent truncated normal random effect. Let $\mathbf{Y}_{n \times k}$ be the observation matrix such that $Y_{ij}$ is the observation for subject $i$ at timepoint $j$.

$$\mathbf{Y}_{n \times k} = \mathbf{X}_{n \times p}\boldsymbol\beta_{p \times k} + t_{n \times 1}\psi_{1 \times k} + \boldsymbol\epsilon_{n \times k}$$

where $X_i$ is the $1 \times p$ vector of covariate values for subject $i$, $\beta_j$ is the $i \times k$ vector of fixed effects coefficients for timepoint $j$, $t_i \stackrel{iid}{\sim} N_{[0,\infty)}(0,1)$ is a truncated normal random effect, $\psi$ is the vector containing skewness parameters for each timepoint, and $\boldsymbol\epsilon_i \sim N_k(0,\boldsymbol\Sigma_{k \times k})$ is the correlated error term.

>>>>

\subsection{Generic Mixture}

The general formulation of the multivariate skew normal (MSN) is as follows.

$$f(\mathbf{y}_i) = \sum_{k = 1}^{K} \pi_{ik} f(\mathbf{y}_i|\boldsymbol\theta_k)$$

@@

We model the effect of covariates on longitudinal development outcomes through the use of a MSN regression model. We adopt a standard stochastic representation of a MSN random variable as the sum of a latent truncated normal random effect and a correlated multivariate normal random error term. More formally, Let $\mathbf{y}_{i}$ be the $J \times 1$ observation vector for subject $i$ such that $y_{ij}$ is the observation for subject $i$ at timepoint $j$. The MSN stochastic representation is as follows.
$$\mathbf{y}_{i} = \boldsymbol\beta^T \mathbf{x}_{i} + t_{i}\boldsymbol\psi + \boldsymbol\epsilon_{i}$$

where $\mathbf{x}_i$ is the $p \times 1$ vector of covariate values for subject $i$, $\boldsymbol\beta$ is the $p \times k$ vector of fixed effects coefficients , $t_i \stackrel{iid}{\sim} N_{[0,\infty)}(0,1)$ is a conditional truncated normal random effect, $\boldsymbol\psi$ is the vector containing cluster-specific skewness parameters, and $\boldsymbol\epsilon_i \sim N_k(0,\boldsymbol\Sigma_{k \times k})$ is the correlated error term. In can be shown that under these conditions $\mathbf{y}_i$ is marginally multivariate skew normal.
